---
title: "WIS 2323: WHERE are Tropical Rain Forests"
author: "Emilio M. Bruna"
date: "updated: `r Sys.Date()`"
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[R]{last update `r Sys.Date()`}
- \fancyfoot[C]{ }
- \fancyhead[R]{WHERE are LTRF, p. \thepage}
- \fancyhead[L]{WIS 2323 - Rain Forests (`r format(Sys.time(), '%Y')`)}
- \usepackage[default]{sourcesanspro}
- \usepackage{parskip}
- \usepackage{xcolor}
- \AtBeginDocument{\let\maketitle\relax}
- \usepackage{sectsty} \sectionfont{\centering}
- \usepackage{titlesec}
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
- \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```
# WIS 2323: WHAT are Tropical Rain Forests


## Objectives and Competencies:   

Learn some of the geographic and climatological factors defining the tropics

## Pre-class Preparation - Instructor

- [ ] Email Students 
- [ ] Bring to Class
> - [ ] Roster
> - [ ] Syllabi
> - [ ] Name cards & markers

## Pre-class Preparation (Students):




### Readings/Videos: 


weather vs. climate
[climate in GNV](https://weatherspark.com/y/16986/Average-Weather-in-Gainesville-Florida-United-States-Year-Round)

```{r timer, echo=FALSE}

Block1_intro <- 30
Break <- 5
trivia <- 60

total_time <- 
  Block1_intro + 
  Break + 
  trivia


time_remain <- 150 - total_time
```
# Class Outline (Total Time: `r total_time` min)
\vspace{0.5cm}

### **Intro: ** (`r Block1_intro` min)

1. Announcements, Movie Project Info

1. Officially: Found between the Tropics of Cancer and Capricorn, but...

2. Rain: At least 2 m (79 inches) and as much as 10 m (394 inches) of rain per year.

3. Temperature: Average daily temperatures range from 20°C (68°F) to 25°C (77°F).

* Note: There is seasonality...but not in temperature. It's seasonality of rain fall.
Offically there are actually rain forests, moist forests, et5c. The ammount of rain, seasonality of rain all determine this, and as a result also have some trees that are [deciduous](https://www.flickr.com/photos/cifor/50480154283/). Here is a better [close-up](here (https://www.flickr.com/photos/cifor/49666688962/)

4. Tall canopy of 45-55 m with occasional "emergent" trees of 75-80 m
* Century Tower: is 48 m (157 ft)
* Tallest tree: detected with Light Detection and Ranging (LiDAR) survey in the Danum Valley Conservation Area (DVCA) in Sabah.  It was measured by climber Unding Jami with a tape measure as being 100.8 meters tall. 


# Useful Reading & Resources
https://www.mapcustomizer.com/

Cities: 
Buffalo, NY (A: 4, M: 39, B: 156)
Cedar Rapids, Iowa (A: 8 M: 14  B: 160 )
Memphis , TN (A: 24  M: 36 B: 254 )
El Paso, TX (A: 16 M: 55 B: 327)
Gainesville, FL (A: 22 M: 32 B: 207)
Cuernavaca, MEX (A: 20 M: 30 B: 227)
Heredia Province, Sarapiquí, Costa Rica (A: 34, M: 33, B: 240)
Gamboa, Panama (A: 32  M: 55 B: 287)
Medellin, Colombia (A: 18 M: 32 B: 309 )
Manaus, Brazil (A: 59 M: 136 B: 325 )
Cuiaba, Brazil (A: 16 M: 23 B: 246 I: 379)
Rio de Janeiro, Brazil (A: 125 M: 126 B: 589)
Cordoba, Argentina (A: 31 M: 59 B: 338)
Montevideo Uruguay (A: 13 M: 24 B: 214)


Yukon Territory, CA A: 3, m: 53, B: 253 (186,272 mi²)
North Dakota, USA A: 10, B: 328, M: 59 (70,704 mi²)
Ohio, USA A: 11, B: 141 (44,825 mi²)
Florida, USA A: 65, B: 576, M: 90 (65,758 mi²)
Puebla, Mex A: 61 B:467 M: 112  (13,224 mi²)
Honduras A: 122, B: 692, M: 166 (43,433 mi²)
Ecuador A: 475 B: 1526, M: 299 (109,483 mi²)
rondonia, Brl B: 639 A: 132 M 130 (91,729 mi²)
Rio de Janeiro A: 125, B: 589, M: 126 (16,871.2 sq mi)
Uruguay A: 37, B: 442, M: 82 (68,037 mi²)


**LIDAR SCAN**
<div class="sketchfab-embed-wrapper"> <iframe title="TLS scan of Menara" frameborder="0" allowfullscreen mozallowfullscreen="true" webkitallowfullscreen="true" allow="fullscreen; autoplay; vr" xr-spatial-tracking execution-while-out-of-viewport execution-while-not-rendered web-share src="https://sketchfab.com/models/03a6159372294c5288f832f7fe1d2fda/embed"> </iframe> <p style="font-size: 13px; font-weight: normal; margin: 5px; color: #4A4A4A;"> <a href="https://sketchfab.com/3d-models/tls-scan-of-menara-03a6159372294c5288f832f7fe1d2fda?utm_medium=embed&utm_campaign=share-popup&utm_content=03a6159372294c5288f832f7fe1d2fda" target="_blank" style="font-weight: bold; color: #1CAAD9;"> TLS scan of Menara </a> by <a href="https://sketchfab.com/ashenkin?utm_medium=embed&utm_campaign=share-popup&utm_content=03a6159372294c5288f832f7fe1d2fda" target="_blank" style="font-weight: bold; color: #1CAAD9;"> Alexander Shenkin </a> on <a href="https://sketchfab.com?utm_medium=embed&utm_campaign=share-popup&utm_content=03a6159372294c5288f832f7fe1d2fda" target="_blank" style="font-weight: bold; color: #1CAAD9;">Sketchfab</a></p></div>


**LIDAR+UAV SCAN**
<div class="sketchfab-embed-wrapper"> <iframe title="LiDAR + UAV model of Menara" frameborder="0" allowfullscreen mozallowfullscreen="true" webkitallowfullscreen="true" allow="fullscreen; autoplay; vr" xr-spatial-tracking execution-while-out-of-viewport execution-while-not-rendered web-share src="https://sketchfab.com/models/b828f4569ab14188b07c9609511330b7/embed"> </iframe> <p style="font-size: 13px; font-weight: normal; margin: 5px; color: #4A4A4A;"> <a href="https://sketchfab.com/3d-models/lidar-uav-model-of-menara-b828f4569ab14188b07c9609511330b7?utm_medium=embed&utm_campaign=share-popup&utm_content=b828f4569ab14188b07c9609511330b7" target="_blank" style="font-weight: bold; color: #1CAAD9;"> LiDAR + UAV model of Menara </a> by <a href="https://sketchfab.com/ashenkin?utm_medium=embed&utm_campaign=share-popup&utm_content=b828f4569ab14188b07c9609511330b7" target="_blank" style="font-weight: bold; color: #1CAAD9;"> Alexander Shenkin </a> on <a href="https://sketchfab.com?utm_medium=embed&utm_campaign=share-popup&utm_content=b828f4569ab14188b07c9609511330b7" target="_blank" style="font-weight: bold; color: #1CAAD9;">Sketchfab</a></p></div>

[The canopy at the Forest Research Institute Malaysia with Crown Shyness](https://commons.wikimedia.org/w/index.php?curid=5188293)(Photo by Mikenorton; Own work, CC BY-SA 3.0).



<!-- NOTE BATS in mammals! -->




<!-- [Forest Height Map](https://eoimages.gsfc.nasa.gov/images/imagerecords/77000/77637/forestheight_ice_2005_lrg.jpg)  -->
<!-- Bonus: Try this video of a [forest in Costa Rica](https://www.youtube.com/watch?v=AhISnxlonxI&t=9s), complete with the dulcet tones of Richard Attenborough's narration.  -->



<!-- 9.  What do you think limits their distribution? (put another way, why don't we have Tropical Rain Forests in Gainesville? Or Michigan? Or London?) -->

<!-- In a rain forest, is rain going to be limited? -->
<!-- For these big trees, is light the limiting factor? What is? -->
<!-- What about for understory plants?  -->



### **Break**  (`r Break` min)    

### **NEXT PART**  (`r trivia` min)      
<!-- UPDATES:  -->

<!-- Each team needs to come up with names OR assign names with each table getting a [table stand with photo](https://www.webstaurantstore.com/choice-12-black-menu-card-holder/176CH12BK.html)  -->

<!-- Set up system in advance to enter scores as we go, or enter responses and tell us if they are correct -->

<!-- Be sure to take attendance to give people points OR ask them to submit their team name afterwards for 20 points -->



\vspace{1.5cm}
# Additional Resources and Reading Materials
\vspace{0.5cm}
