---
title: "WIS 2323 - \"Analytical Essay"
author: "Emilio M. Bruna"
date: "`r format(Sys.time(), '%Y')`"
# output:
#   html_document:
#     df_print: paged
geometry: margin=1in
fontsize: 12pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot{}
- \fancyhead[R]{Analytic Essay Assignment, p. \thepage}
- "\\fancyhead[L]{WIS 2323 - Rain Forests (`r format(Sys.time(), '%Y')`)}"
- \usepackage[default]{sourcesanspro}
- \usepackage{parskip}
- \usepackage{xcolor}
- \AtBeginDocument{\let\maketitle\relax}
- \usepackage{titlesec}
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
- \usepackage{sectsty} \subsectionfont{\centering\color{darkmidnightblue}}
- \usepackage{sectsty} \subsubsectionfont{\centering\color{black}}
- \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
# output: html_document
  # editor_options:
  # markdown: null
  # wrap: 72
mainfont: SourceSansPro
# knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding,
#   output_dir = "./static/uploads") })
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```


<!-- TODO UPDATES:  -->

<!-- https://news.mongabay.com/2022/03/study-indonesias-forest-clearing-moratorium-underdelivered-but-so-did-donors/ -->
<!-- https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0266178 -->

# WIS 2323: Analytical Essay
<!-- \vspace{0.3cm} -->
### Due: 28 October 2025 | Value: 150 Points | Submission: Canvas

## Assignment Overview  

Brazil and Indonesia are among the world's countries with the most tropical forest (ranked 1st and 3rd, respectively). Considerable attention is therefore devoted to documenting the amount of forest cover in both countries, as well as investigating the factors responsible changes in forest cover over time. These factors can be both internal (i.e., domestic social, economic, and political issues) and external (i.e., international events, foreign policy). **_Your assignment is to analyze Brazil and Indonesia's forest cover from 2002-2024 and suggest internal or external factors could be responsible for any major changes the rate of forest loss._**

**To complete this analytic essay you must:** 

1. **Download data:** Use the Global Forest Watch Database ([https://www.globalforestwatch.org/](https://www.globalforestwatch.org/)) to find and download data for the amount of tree cover and tree cover loss in Brazil and Indonesia between 2002-2024.

2. **Calculate the following:**

  > a) the *total* tree cover lost from 2002 to 2024, 
  > b) the *percentage* of 2002 tree cover remaining in 2024,
  > c) the *annual amount of forest lost* between 2002 and 2024, and
  > d) the *annual rate of forest loss* between 2002 to 2024.
  
  > ***Note:*** _Use km^2^ for your analyses, figures, and report (you will have to convert from hectares to km^2^)._ 

3. **Prepare two graphs:** the first with the data from (2c), and the second with the data from (2d). Each plot should be a _line graph_ and include the data for both countries. Make sure each graph has properly labeled axes, a legend, and a title.

4. **Analyze and investigate the patterns of forest loss:** Once you have graphed the data, see if you can identify any trends in the rate of forest loss over time. Pay particular attention to trends before and after 2016. Then conduct research to identify any factors that might be responsible for the  observed trends. These could be within-country factors (e.g., government policies or national economic expansion or contraction) or international events (e.g., implementation of forest preservation treaties, international trade agreements or disputes).

5. **Write an essay** in which you summarize and compare the rates of deforestation in Brazil and Indonesia, and present your conclusions regarding the mechanisms that are likely responsible for the trends you graphed. The essay should be $\geq$ 1000-1200 words (1000 words is approx 2-2.5 pages of single-spaced text). Note that you should *not* respond to each questions with bullet points - the responses should be in essay format, using carefully revised writing, and thorough descriptions of the observed patterns followed by their interpretation. You may wish to consider how your conclusions about patterns in forest loss change if considering longer time intervals (i.e., 10-20 years) vs. shorter ones (1-5 years). You might also try and predict the amount of forest loss going forward.

  > The essay must include a bibliography with citations of at least 4 relevant sources that support your conclusions. The sources should be credible and meet the standard for academic research - ideally _Primary Sources_ (see below for more guidance). The bibliography can be formatted in any standard style  (i.e, MLA, APA, Chicago, etc). 

6. **Submission:** Your essay with bibliography and 2 graphs should be submitted via canvas as a single pdf document. The essay should be single-spaced text with 1 inch margins. Label the figures as Figure 1 and Figure 2. The essay is **due 2 November 2023** and has a **value = 150 Points**. It will be graded with the rubric below.


\vspace{1cm}

## Additional Information


1. **A word about Sources: **The UF Smathers Libraries are among the best academic libraries in the country, and the librarians are *phenomenal*. Ask them for help both with looking for information for your papers, for how to determine if something is an acceptable source of information for a paper (see https://guides.uflib.ufl.edu/libraryresearch/sources and https://guides.uflib.ufl.edu/libraryresearch/evaluatingsources). Or better still: go to Marston Science Library and ask one of our outstanding librarians to help you track down sources), and for how to evaluate the quality of the information presented. They have an online tutorial on [Library Research Basics](https://guides.uflib.ufl.edu/libraryresearch/libraryresearch) but you can actually get help with class assignments from the [Research Assistance Desk](https://librarywest.uflib.ufl.edu/services-resources/research-assistance-desk/) (both walk-ups and appointments are available). Take advantage of this paper to get to know the Librarians and the resources we have to make your life easier.

2. **A word about Bibliographies: **There are few things in life worse than typing up bibliographies for papers. It takes forever and you will make tons of mistakes. Worse, the next professor will require the bibliography in a different format, so even if you can re-use a reference you have to retype it. If only there was a way to automate the process so you could print up a bibliography in a few seconds at the touch of a button. Behold! I give you [Endnote](https://guides.uflib.ufl.edu/endnote). Or [Zotero](https://guides.uflib.ufl.edu/zoterocite). Take your pick, both are free for UF students. Never type a bibliography again. Need help learning to use them? Ask me in office hours or ask the librarian listed of the page of the one you're trying to learn. 


```{=tex}
\newpage
\blandscape
```
## RUBRIC
<!-- table row spacing set with arraystretch -->
<!-- \renewcommand{\arraystretch}{0.75}  -->
```{r rubric, echo=FALSE, message = FALSE, warning=FALSE, results='asis'}
# rubric <- read_csv("./class_materials/projects/analytical_essay_deforestation/rubric.csv")
rubric <- read_csv("./course_projects/analytical_essay_deforestation/rubric.csv") %>% 
  replace_na(list("Criterion" = " ")) %>% 
  mutate(`Excellent - Very Good`=paste(`Excellent - Very Good`," (",floor(e1),"-",floor(e2)," pts)",sep="")) %>% 
  mutate(`Good`=paste(`Good`," (",floor(g1),"-",floor(g2)," pts)",sep="")) %>% 
  mutate(`Satisfactory`=paste(`Satisfactory`," (",floor(s1)-1,"-",floor(s2)," pts)",sep="")) %>% 
  mutate(`Unsatisfactory`=paste(`Unsatisfactory`," (< ",floor(s1)-1," pts)",sep="")) %>% 
  select(-c(e1,e2,g1,g2,s1,s2))

variables<-names(rubric)

tmp<-knitr::kable(rubric,
     align="l",
     "latex",
     # caption.short = "Demo Table",
     # escape=FALSE,
     # row.names = FALSE,
     # caption = "--",
     booktabs=T,
     linesep = "" #removes the blank line after every 5 lines
     ) %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = T, 
                latex_options = c("scale_down",
                                  "repeat_header"),
               font_size = 10,
               position = "left") %>%
   column_spec(1, width = "8em") %>% 
   column_spec(2, width = "16em") %>% 
   column_spec(3, width = "16em") %>% 
   column_spec(4, width = "16em") %>% 
  column_spec(5, width = "16em") %>% 
  row_spec(c(1:4), hline_after = T) %>% 
   row_spec(c(0), bold = T)

knitr::asis_output(tmp)
```


```{=tex}
\elandscape
```






