---
title: "WIS 2323: ------"
date: "updated: `r Sys.Date()`"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot{}
- \fancyhead[R]{Forest Cover & Loss, p. \thepage}
- \fancyhead[L]{WIS 2323 - Rain Forests (`r format(Sys.time(), '%Y')`)}
- \usepackage[default]{sourcesanspro}
- \usepackage{parskip}
- \usepackage{xcolor}
- \AtBeginDocument{\let\maketitle\relax}
# - \usepackage{sectsty} \allsectionsfont{\raggedright}
- \usepackage{sectsty} \sectionfont{\centering}
- \usepackage{titlesec}
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
- \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

## Objectives and Competencies:   

1.To learn how different organizations define 'forest'
2. Be able to compare FRA and GFW criteria  
3. Gain an overview of how much forest there is in the world, and how much of this is in the tropics
4. Learn how the drivers of deforestation differ between different tropical regions.
5. Learn the importance of roads and learn to recognize the fishbone deforestation pattern


## Pre-Class Preparation - Instructor:  

- [ ] Email Reminder to read news paper article by S Nolen
- [ ] **Bring to Class:**  


## Pre-class Preparation - Students:


### **Reading or Watch**

## Class Outline


```{r timer, echo=FALSE}

Block1_intro <- 25
Block1_breakout <- 30
Break <- 10
return_results <- 20
questions <- 10
intro_next_session <- 10
  
  
total_time <- 
  Block1_intro + 
  Block1_breakout + 
  Break + 
  return_results + 
  questions +
  intro_next_session 


time_remain <- 50 - total_time
```



<!-- ### **Topic 1 Overview: Paperless Data Collection** (`r t1_overview` min) -->

### **Topic 1: Patterns of Forest Cover & Consequences of Forest Loss

1) Forest - Where is it, How much of it, & The importance of Intact Forest
2) Primary Drivers of Forest Loss
3) What happens to a landscape when you clear forest?
4) What happens when you abandon these areas?


1) Forest has a [technical definition](https://unstats.un.org/sdgs/metadata/files/Metadata-15-01-01.pdf) according to the UN FAO. Actually, there are many [technical defintions](https://www.iucn.org/sites/dev/files/content/documents/primary-forest-report-final-march-2021.pdf). Plus two of the main resources - FAO Forest Resources Assessment (FRA) and WRI Global Forest Watch (GFW) [differ in some significant ways](https://www.wri.org/insights/insider-global-forest-watch-and-forest-resources-assessment-explained-5-graphics).  

### Demo

1. Click "Map" on the main menu bar.
2. "Land Cover" on the Left icon menu
3. x-out the layers you don't need on the legend

<!-- There are `r time_remain` min remaining that can be used to ----- -->

# Additional Resources and Reading Mateirals

2.
3.

