---
title: "IDS 2935: Patterns of Biodiversity"
author: "Emilio M. Bruna"
date: "updated: `r Sys.Date()`"
mainfont: SourceSansPro
geometry: margin=1in
fontsize: 12pt
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[R]{last update `r Sys.Date()`}
- \fancyfoot[C]{ }
- \fancyhead[R]{Patterns of Biodiversity, p. \thepage}
- \fancyhead[L]{IDS 2935 - Rain Forests (`r format(Sys.time(), '%Y')`)}
- \usepackage[default]{sourcesanspro}
- \usepackage{parskip}
- \usepackage{xcolor}
- \AtBeginDocument{\let\maketitle\relax}
- \usepackage{sectsty} \sectionfont{\centering}
- \usepackage{titlesec}
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
- \titlespacing{\section}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \titlespacing{\subsubsection}{0pt}{12pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
output:
  pdf_document: default
  html_document:
    df_print: paged
  rmdformats::robobook: default
editor_options: 
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```
# IDS 2935: Patterns of Biodiversity


## Objectives and Competencies:   

1. Learn the difference between species richness and species diversity 
2. Learn and be able to describe two major patterns of species richness related to the tropics

> (A) Latitudinal Diversity Gradient
> (B) High diversity in a site, but few common species (but most are very rare)

3. Learn to use iNaturalist to find or record biodiversity data
4. Download data, import into google docs, plot, and analyze


## Pre-class Preparation - Instructor

- [ ] Email Students  
> - [ ] class location
> - [ ] reminders

## Bring to Class

- [ ] Roster
- [ ] Syllabus
- [ ] Name cards & markers


## Pre-class Preparation - Students


### Readings/Videos: 



```{r timer, echo=FALSE}

Block1_intro <- 30
Break <- 5
trivia <- 60

total_time <- 
  Block1_intro + 
  Break + 
  trivia


time_remain <- 150 - total_time
```
# Class Outline (Total Time: `r total_time` min)
\vspace{0.5cm}

### **Intro: ** (`r Block1_intro` min)

1. Announcements:

1. Reminder: seasonality...but not in temperature, it's seasonality of rain fall.

2. Tallest tree: detected with Light Detection and Ranging (LiDAR) survey in the Danum Valley Conservation Area (DVCA) in Sabah.  It was measured by climber Unding Jami with a tape measure as being 100.8 meters tall. 

### Global Biodiversity

1. How many species?
2. What are they?

Everyone got a different color post-it note, stand up / sit down to show the relative number of species in different groups.


## Latitudinal Diversity Gradient

Cities: 
58: Juneau, AK (A: 3, M: 31, B: 157)
49: Winnipeg, Manitoba, CA  (A: 6, M: 35, B: 248)
43: Buffalo, NY (A: 4, M: 39, B: 156)
41: Cedar Rapids, Iowa (A: 8 M: 14  B: 160 )
<!-- Memphis , TN (A: 24  M: 36 B: 254 ) -->
36: Nashville, TN (A: 24  M: 44 B: 274 ) *Search for "Greater Nashville) 
31: El Paso, TX (A: 16 M: 55 B: 327)
29: Gainesville, FL (A: 22 M: 32 B: 207)
25: Monterrey, MEX (A: 20 M: 67 B: 376)
18: Cuernavaca, MEX (A: 19 M: 31 B: 235)
10: Heredia Province, Sarapiquí, Costa Rica (A: 34, M: 33, B: 240)
9: Gamboa, Panama (A: 32  M: 55 B: 287)
6: Medellin, Colombia (A: 18 M: 32 B: 309 )
-3: Manaus, Brazil (A: 59 M: 136 B: 325 )
-10: Rio Branco, AC, BR (A: 51 M: 45 B: 348 )
-15: Cuiaba, Brazil (A: 16 M: 23 B: 246 I: 379)
-22: Rio de Janeiro, Brazil (A: 125 M: 126 B: 589)
-25: Curitiba, BR (A: 44 M: 55 B: 324)  (greater met area)
<!-- Porto Alegre (A: 11 M: 20 B: 165) -->
<!-- Asuncion, Paraguay (A: 10, M: 6, B: 131) -->
<!-- -31 Cordoba, Argentina (A: 31 M: 59 B: 338) -->
-32: Rosario, Argentina (A: 12 M: 22 B: 244)
-34: Montevideo Uruguay (A: 13 M: 24 B: 214)
-39: Valdivia, Chile (county) (A: 18 M: 45 B: 150)
-51: Torres del Paine, Chile (Open Space) (A: 1 M: 14 B: 90)
-53: Punta Arenas Chile (A: 0, M: 9, B: 80)


Verifiable, Research Grade
Yukon Territory, CA A: 3, m: 53, B: 253 (186,272 mi²)
North Dakota, USA A: 10, B: 328, M: 59 (70,704 mi²)
Ohio, USA A: 424, B: 141 M: 65 (44,825 mi²)
Florida, USA A: 65, B: 576, M: 90 (65,758 mi²)
Puebla, Mex A: 61 B:467 M: 112  (13,224 mi²)
Honduras A: 116, B: 691, M: 154 (43,433 mi²)
Ecuador A: 475 B: 1467, M: 299 (109,483 mi²)
Rondonia, Brl B: 639 A: 132 M 130 (91,729 mi²)
Rio de Janeiro A: 125, B: 589, M: 126 (16,871.2 sq mi)
Uruguay A: 37, B: 442, M: 82 (68,037 mi²)



<!-- NOTE BATS in mammals! -->




<!-- [Forest Height Map](https://eoimages.gsfc.nasa.gov/images/imagerecords/77000/77637/forestheight_ice_2005_lrg.jpg)  -->
<!-- Bonus: Try this video of a [forest in Costa Rica](https://www.youtube.com/watch?v=AhISnxlonxI&t=9s), complete with the dulcet tones of Richard Attenborough's narration.  -->



<!-- 9.  What do you think limits their distribution? (put another way, why don't we have Tropical Rain Forests in Gainesville? Or Michigan? Or London?) -->

<!-- In a rain forest, is rain going to be limited? -->
<!-- For these big trees, is light the limiting factor? What is? -->
<!-- What about for understory plants?  -->



### **Break**  (`r Break` min)    

### **NEXT PART**  (`r trivia` min)      
<!-- UPDATES:  -->

<!-- Each team needs to come up with names OR assign names with each table getting a [table stand with photo](https://www.webstaurantstore.com/choice-12-black-menu-card-holder/176CH12BK.html)  -->

<!-- Set up system in advance to enter scores as we go, or enter responses and tell us if they are correct -->

<!-- Be sure to take attendance to give people points OR ask them to submit their team name afterwards for 20 points -->



\vspace{1.5cm}
# Additional Resources & Reading
\vspace{0.5cm}

1. Make maps with locations on them: https://www.mapcustomizer.com/

